---
title       : "Fatal Police Shootings: 2015 - 2020"
subtitle    : 
author      : "Clay Glad"
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

<br>  
The <a href = "https://washingtonpost.com">Washington Post</a> has tracked all 
fatal police shootings in the US since 2015.  
On this page you can display this data on a map, filtering for race, age, and
sex.  
<br>  
If you click down on the map to the level of individual shootings, you will see  
the date and location of the shooting, as well as the name of the victim.  
<br>  
The pie chart beneath the map displays what fraction of all fatal shootings were  
of those victims in the selected filter.  
<br>  
Reference: <a href= "https://www.washingtonpost.com/graphics/investigations/police-shootings-database/">Washington Post: Fatal Force</a>  
Data: <a href = "https://github.com/washingtonpost/data-police-shootings">Washington Post GitHub repository on fatal police shootings</a>

---
<br><br>  
The program is written in Shiny R.  
<br>  
Leaflet was used to generate the maps and Plotly to plot the pie charts.  
The tidyverse, data.table, and scales libraries were also used.  
<br>  
Filtering the data was achieved by first factorizing the ages into seven  
distinct categories (under 13, 13-19, 20-29, 30-39, 40-49, 50-59, 60 and  
over 60). Then the data was twice subsetted: one data set for map generation  
and the other for plot generation. The code for this appears on the next slide.


---
```{r chunk1, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(data.table)
library(scales)
library(shiny)
library(shinythemes)
library(leaflet)
library(plotly)
```
```{r chunk2, message = FALSE, warning = FALSE}
shootings = fread("fatal-police-shootings-data.csv")
shootings = shootings %>%
        rename(sex = gender) %>%
        mutate(age = case_when(
                age < 13 ~ "kid",
                age > 12 & age < 20 ~ "teen",
                age > 19 & age < 30 ~ "twenties",
                age > 29 & age < 40 ~ "thirties",
                age > 39 & age < 50 ~ "forties",
                age > 49 & age < 60 ~ "fifties",
                age >= 60 ~ "senior"))

## Different subsetted data tables for generating the map and the plot

mapShootings = shootings %>%
        unite("cityState", city:state, sep = ", ") %>%
        select(name, date, age, sex, race, cityState, longitude, 
               latitude)

plotShootings = shootings %>%
        select(age, sex, race)
```

--- 

<br><br>  
The Shiny UI was achieved by using radio buttons for the filtering factors.  
<br> 
The main panel takes three outputs: the Leaflet-generated map, text determined  
by both values from the data set and canned responses, and finally the  
Plotly-generated pie chart.  
<br>  
If no filters are selected, no plot is generated and the explantory text  
that appears on first loading is reloaded.